---
title: "Data_wrangling_II"
output: github_document
---

```{r}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)
library(httr)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
New packages: httr and rvest, have to load seperately, part of tidy verse ecosystem tho


##scrapping table
Want the first tabel from [this page]: http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.html

reading the html
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"


drug_use_html <- read_html(url)
```
Etratcing the tables: selector gadgets;
```{r}
tabl_marj <- 
drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
  
```
^A total of 15 tables are extratced.
we are going to fcus on the first tabel (first(), nth(), last());
slice used to access specifc data lines, (-1) removed first line, as_tibble turns dataframe int a tibble


##star wars 
want data from here: 

```{r}

swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```
Now we only grab elements we want, but for this page we dont know what css tag IMDB uses for movie info pages. This is where selector gagets comes in 
```{r}
title_vec <- 
  swm_html %>% 
  html_nodes(css = ".lister-item-header a") %>% 
  html_text()


gross_rev_vec <- 
   swm_html %>% 
  html_nodes(css = ".text-muted .ghost~ .text-muted+ span") %>% 
  html_text()

run_time_vec <- 
  swm_html %>% 
  html_nodes(css = ".runtime") %>% 
  html_text()

swm_df <- 
  tibble(
    tittle  = title_vec,
    gross_rev = gross_rev_vec,
    runtime = run_time_vec)
```

