---
title: "Data_wrangling_II"
output: github_document
---

```{r}
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)
library(httr)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
New packages: httr and rvest, have to load seperately, part of tidy verse ecosystem tho


##scrapping table
Want the first tabel from [this page]: http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.html

reading the html
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"


drug_use_html <- read_html(url)
```
Etratcing the tables: selector gadgets;
```{r}
tabl_marj <- 
drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
  
```
^A total of 15 tables are extratced.
we are going to fcus on the first tabel (first(), nth(), last());
slice used to access specifc data lines, (-1) removed first line, as_tibble turns dataframe int a tibble


##star wars 
want data from here: 

```{r}

swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```
Now we only grab elements we want, but for this page we dont know what css tag IMDB uses for movie info pages. This is where selector gagets comes in 
```{r}
title_vec <- 
  swm_html %>% 
  html_nodes(css = ".lister-item-header a") %>% 
  html_text()


gross_rev_vec <- 
   swm_html %>% 
  html_nodes(css = ".text-muted .ghost~ .text-muted+ span") %>% 
  html_text()

run_time_vec <- 
  swm_html %>% 
  html_nodes(css = ".runtime") %>% 
  html_text()

swm_df <- 
  tibble(
    tittle  = title_vec,
    gross_rev = gross_rev_vec,
    runtime = run_time_vec)
```


API's 
make a reqeust to a API ot get data
##NYC water

```{r}
nyc_water <- 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")


nyc_water <- 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()
```
same process different data:
```{r}
brfss_2010 <- 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv", 
      query = list("$limit" = 5000)) %>% 
  content("parsed")
```


##some data are as nice as these exapmles 
Pokemon example 

```{r}

pokemon_data <- 
  GET("https://pokeapi.co/api/v2/pokemon/1") %>% 
  content()
  

pokemon_data$name
pokemon_data$height
```






##Strings and factors 


```{r}
library(p8105.datasets)
```
Matching a pattern
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_replace(string_vec, "jeff", "Jeff")
```
Exact macth 
```{r}

string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

```
Designate a list of characters for matching 
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )
str_detect(string_vec,"[Bb]ush")
```

or a range of characters
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```
The character . matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )
str_detect(string_vec, "7.11")
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird. Thats why we use "\\"to specific that the charater follows is the exact one we are looking for 
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```
##Factors
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
as.numeric(vec_sex)

vec_sex = fct_relevel(vec_sex, "male")
vec_sex
as.numeric(vec_sex)
```
^there is also fct_reorder, fct_recode, etc.


```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)

data_marj <-  
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

another example
```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))
```


```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```
better way to do it:
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)


rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
fct_infreq: by number of observations with each level (largest first)
fct_inorder(): by the order in which they first appear.
fct_inseq(): by numeric value of level.

suppose we want tp rename a borough


```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
This however changed the bror variable back to strings 
using base r wont work either becuase fatcors need to bne specific values, chaning the name to something tah doesnt eost wont work

this si when we use fct_recode:
```{r}
rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

